### 2017-03-26 paper/oopsla17-submission/evaluation-accuracy.tex
feature selection

\paragraph{Our Classifiers}
We evaluate five classifiers, each trained on the full set of 276
features: 44 local syntactic features, 176 contextual syntactic
features, 55 typing features, and a single expression size feature.
%
\ES{should explain the make-up of these groups}
%
We preemptively discard expressions that are not part of the minimal
type error slice, we will explain the rationale for this in
\autoref{sec:feature-utility}.
%
Our classifiers are:
%
\begin{enumerate}
\item A logistic regression trained with a learning rate
  $\alpha = 0.001$, an L2 regularization rate $\lambda = 0.001$, and a
  mini-batch size of 200.
\item A decision tree trained with the \ES{HUMA?} algorithm and \ES{HUMA: hyperparams?}.
\item A random forest trained with the \ES{HUMA?} algorithm and 30 estimators \ES{HUMA: hyperparams?}.
\item Two multi-layer perceptrons, both trained with $\alpha = 0.001$,
  $\lambda = 0.001$, and a mini-batch size of 200. The first MLP
  contains a single hidden layer of 10 neurons, and the second contains
  a hidden layer of 500 neurons. This allows us to investigate how well
  the MLP can \emph{compress} its model.
\end{enumerate}
%
All classifiers were trained for 20 epochs on one dataset
--- \ie they were shown each program 20 times ---
before being evaluated on the other.

@@

Surprisingly, there is little variation in accuracy between classifiers
with the exception of the \linear model, they all achieve around 70\%
Top-1 accuracy and around 90\% Top-3 accuracy.
%
This would appear to suggest that the model they learn is relatively
simple.
%
In particular, notice that although the \hiddenT has $50\times$ \emph{fewer}
hidden features than the \hiddenFH, it only loses around 2\% accuracy.
% In particular, notice that the \hiddenT only loses around 2\% accuracy
% compared to the \hiddenFH,
%
We also note that our classifiers consistently perform better when
trained on the \FALL programs and tested on the \SPRING programs than
vice versa, they appear to generalize better from the \FALL data.

>>>>

\paragraph{Our Classifiers}
We evaluate five classifiers, each trained on the full set of features.
% features: 44 local syntactic features, 176 contextual syntactic
% features, 55 typing features, and a single expression size feature.
% %
% \ES{should explain the make-up of these groups}
%
We preemptively discard expressions that are not part of the minimal
type error slice --- we will explain the rationale for this in
\autoref{sec:feature-utility} --- and thus the final feature count is
276.
%
Our classifiers are:
%
\begin{enumerate}
\item A logistic regression trained with a learning rate
  $\alpha = 0.001$, an L2 regularization rate $\lambda = 0.001$, and a
  mini-batch size of 200.
\item A decision tree trained with the \ES{HUMA?} algorithm and \ES{HUMA: hyperparams?}.
\item A random forest trained with the \ES{HUMA?} algorithm and 30 estimators \ES{HUMA: hyperparams?}.
\item Two multi-layer perceptrons, both trained with $\alpha = 0.001$,
  $\lambda = 0.001$, and a mini-batch size of 200. The first MLP
  contains a single hidden layer of 10 neurons, and the second contains
  a hidden layer of 500 neurons. This allows us to investigate how well
  the MLP can \emph{compress} its model.
\end{enumerate}
%
All classifiers were trained for 20 epochs on one dataset
--- \ie they were shown each program 20 times ---
before being evaluated on the other.

@@

Surprisingly, there is little variation in accuracy between classifiers
with the exception of the \linear model, they all achieve around 70\%
Top-1 accuracy and around 90\% Top-3 accuracy.
%
This would appear to suggest that the model they learn is relatively
simple.
%
In particular, notice that although the \hiddenT has $50\times$ \emph{fewer}
hidden neurons than the \hiddenFH, it only loses around 2\% accuracy.
% In particular, notice that the \hiddenT only loses around 2\% accuracy
% compared to the \hiddenFH,
%
We also note that our classifiers consistently perform better when
trained on the \FALL programs and tested on the \SPRING programs than
vice versa, they appear to generalize better from the \FALL data.

### 2017-03-26 paper/oopsla17-submission/evaluation-utility.tex
feature selection

Next, we will investigate the relative impact of the other three classes
of features discussed in \autoref{sec:features}.
%
For this we consider again a baseline of only local syntactic features,
extended by each combination of
%
(1) expression size,
(2) contextual syntactic features, and
(3) typing features.
%
As before we perform a 10-fold cross-validation with $\alpha = 0.001$,
$\lambda = 0.001$, and a mini-batch size of 200, but we
train for a full 20 epochs.

>>>>

Continuing on, we will investigate the relative impact of the other
three classes of features discussed in \autoref{sec:features}, assuming
we have discarded expressions not in the type error slice.
%
For this experiment we consider again a baseline of only local syntactic
features, extended by each combination of
%
(1) expression size,
(2) contextual syntactic features, and
(3) typing features.
%
As before we perform a 10-fold cross-validation with $\alpha = 0.001$,
$\lambda = 0.001$, and a mini-batch size of 200, but we
train for a full 20 epochs.

### 2017-03-26 paper/oopsla17-submission/evaluation.tex
feature selection

We identified the fixes for each ill-typed program with an
expression-level diff~\citep{Lempsink2009-xf}.
%
We consider two sources of changes:
%
\begin{enumerate}
\item If an expression has been removed wholesale, \eg if $\eapp{f}{x}$
  is rewritten to $\eapp{g}{x}$, we will mark the expression $f$ as
  changed, as it has been replaced by $g$.
\item If a new expression has been inserted \emph{around} an existing
  expression, \eg if $\eapp{f}{x}$ is rewritten to
  $\eplus{\eapp{f}{x}}{1}$, we will mark the application expression
  $\eapp{f}{x}$ (but not $f$ or $x$) as changed, as the $+$ operator now
  occupies the original location of the application.
\end{enumerate}
%
As the students may have made many, potentially unrelated, changes
between compilations, we discard any program pairs where more than 40\%
of the sub-expressions have changed.
%
We picked 40\% as an estimate of the inflection point where we could
still retain the large majority of program pairs.

\paragraph{Accuracy Metric}
All of the tools we compare (with the exception of \ocaml) can produce a
list of potential error locations, which might in the limit include the
entire type error slice.
%
However, in a study of fault localization techniques,
\citet{Kochhar2016-oc} show that most developers will not consider more
than around five potential error locations before falling back to manual
debugging.
%
Type errors are relatively simple in comparison to general fault
localization, thus we limit our evaluation to the top three predictions
of each tool.
%
We evaluate each tool on whether a changed expression occurred in its
top 1, top 2, or top 3 predictions.

>>>>

\subsubsection{Feature Selection}
We extract a set of 277 features from each sub-expression in a student
program, including:
%
\begin{enumerate}
\item 44 local syntactic features. In addition to the syntax of \lang,
  we support the full range of arithmetic operators (integer and
  floating point), equality and comparison operators, character and
  string literals, and a user-defined |expr| type of simple arithmetic
  expressions. We will discuss the challenge of supporting other
  user-defined types in \autoref{sec:discussion}.
\item 176 contextual syntactic features. For each sub-expression we
  additionally extract the local syntactic features of its parent and
  first, second, and third (left-to-right) children. If an expression
  does not have a parent or children, these features will simply be
  disabled. If an expression has more than three children, the
  classifiers will be unable to learn anything about the additional
  children.
\item 55 typing features. In addition to the types of \lang, we support
  |int|s, |float|s, |char|s, |string|s, and the user-defined |expr|
  mentioned above. These features are extracted for each sub-expression
  and for the contextual sub-expressions.
\item 1 feature denoting the size of each sub-expression.
\item 1 feature denoting whether each sub-expression is part of the
  minimal type error slice.
\end{enumerate}

\subsubsection{Blame Oracle}
We identified the fixes for each ill-typed program with an
expression-level diff~\citep{Lempsink2009-xf}.
%
We consider two sources of changes:
%
\begin{enumerate}
\item If an expression has been removed wholesale, \eg if $\eapp{f}{x}$
  is rewritten to $\eapp{g}{x}$, we will mark the expression $f$ as
  changed, as it has been replaced by $g$.
\item If a new expression has been inserted \emph{around} an existing
  expression, \eg if $\eapp{f}{x}$ is rewritten to
  $\eplus{\eapp{f}{x}}{1}$, we will mark the application expression
  $\eapp{f}{x}$ (but not $f$ or $x$) as changed, as the $+$ operator now
  occupies the original location of the application.
\end{enumerate}
%
As the students may have made many, potentially unrelated, changes
between compilations, we discard any program pairs where more than 40\%
of the sub-expressions have changed.
%
We picked 40\% as an estimate of the inflection point where we could
still retain the large majority of program pairs.

\subsubsection{Accuracy Metric}
All of the tools we compare (with the exception of \ocaml) can produce a
list of potential error locations, which might in the limit include the
entire type error slice.
%
However, in a study of fault localization techniques,
\citet{Kochhar2016-oc} show that most developers will not consider more
than around five potential error locations before falling back to manual
debugging.
%
Type errors are relatively simple in comparison to general fault
localization, thus we limit our evaluation to the top three predictions
of each tool.
%
We evaluate each tool on whether a changed expression occurred in its
top 1, top 2, or top 3 predictions.

### 2017-03-26 paper/oopsla17-submission/gallery.tex
feature selection

\paragraph{Computing the Fixed Point of a Function}
Finally, our students must write a |fixpoint| function that computes the
fixed point of a given function |f|, starting from a base case |b|.
%
As a hint we first have them write a |wwhile| function that performs the
functional equivalent of a while-loop, repeatedly passing a function's
output back in until it receives a (boolean) signal to stop.
%
\begin{ecode}
  let rec wwhile (f, b) =
    match f b with
    | (x, false) -> x
    | (x, true)  -> wwhile (f, x)
\end{ecode}
\lstset{firstnumber=last}
%
The |fixpoint| function can then be written as a clever instantiation of
the arguments to |wwhile|.
%
\begin{ecode}
  let fixpoint (f, b) =
    let g = __let bb = f b in ___=(bb, (bb = b))=_ in
    wwhile (g, b)
\end{ecode}
\lstset{firstnumber=1}
%
Sadly, our student has forgotten that |g| should itself be a function.
%
As a result, she passes a pair of a \emph{pair} and a starting value to
|wwhile|, rather than a pair of a \emph{function} and a starting value.
%
\sherrloc deduces that the call to |wwhile| is likely correct (\ocaml
actually blames the use of |g|), but identifies the construction of the
pair inside |g| as the most likely culprit, while our Decision Tree
correctly identifies the \emph{definition} of |g| as the source of the
error.
\ES{TODO: WHY?}

>>>>

\paragraph{Computing the Fixed Point of a Function}
Finally, our students must write a |fixpoint| function that computes the
fixed point of a given function |f|, starting from an initial value |b|.
%
As a hint we first have them write a |wwhile| function that performs the
functional equivalent of a while-loop, repeatedly passing a function's
output back in until it receives a (boolean) signal to stop.
%
\begin{ecode}
  let rec wwhile (f, b) =
    match f b with
    | (x, false) -> x
    | (x, true)  -> wwhile (f, x)
\end{ecode}
\lstset{firstnumber=last}
%
The |fixpoint| function can then be written as a clever instantiation of
the arguments to |wwhile|.
%
\begin{ecode}
  let fixpoint (f, b) =
    let g = __let bb = f b in ___=(bb, (bb = b))=_ in
    wwhile (g, b)
\end{ecode}
\lstset{firstnumber=1}
%
Sadly, our student has forgotten that |g| should itself be a function.
%
As a result, she passes a pair of a \emph{pair} and a starting value to
|wwhile|, rather than a pair of a \emph{function} and a starting value.
%
\sherrloc deduces that the call to |wwhile| is likely correct (\ocaml
actually blames the use of |g|), but identifies the construction of the
pair inside |g| as the most likely culprit, while our Decision Tree
correctly identifies the \emph{definition} of |g| as the source of the
error.
\ES{TODO: WHY?}
